---
title: "MSDS 6306 DDS Assignment 11 - Time Series"
author: "Luke Pierce"
date: "April 7, 2018"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Set to your working directory after adding Wand data files
setwd("P:/SMU/MSDS 6306 Doing Data Science/Lecture Assignments/Live Session Unit 11 Assignment")

# Clean the workspace, remove all existing Objects
rm(list=ls())

# Load the needed libraries

library(fpp)
library(fpp2)
library(dplyr)
library("dygraphs")
library("forecast")
library("xts")
library(knitr)
library(kableExtra)

# Set RMarkdown plots to a default 12 by 8 inches
knitr::opts_chunk$set(fig.width=10, fig.height=6)

```

####GitHub Repository: https://github.com/lp5510/Homework/tree/master/LiveSession11

# Questions

###1. Warm Up: Brief Financial Data 

####a. Natively in R, you have access to sample data sets of prominent stocks over time. We'll be using EuStockMarkets for this question. Type help(EuStockMarkets) to learn more. From these data, pull specifically the DAX index. For all questions in this assignment, you're welcome to normalize (or don't!) how you see fit, but, if you choose to, please document what you're doing and why for the grader. It's not necessary for the purpose of this assignment.

```{r Read EuStockMarkets and extract DAX data 1991 - 1999 P1, echo = TRUE}

# Read EuStockMarkets data 1991 - 1999 
EUdata <- window(EuStockMarkets)

# Diagnostics
str(EUdata)

```

```{r Read EuStockMarkets and extract DAX data 1991 - 1999 P2, echo = TRUE}

head(EUdata)

```


```{r Extract DAX data 1991 - 1999, echo = TRUE}

# Extract DAX data only
DAXdata <- EUdata[,1]

```

####b. These are annual European Stock Data from 1990 onward. Create a rudimentary plot of the data. Make the line blue. Give an informative title. Label the axes accurately. In 1997, an event happened you want to indicate; add a vertical red line to your plot which divides pre-1997 and post-1997 information.

```{r Plot DAX time series data, echo = TRUE}

# Plot DAX time series data, annotate 1997
plot(DAXdata, ylab="Closing Price", xlab="Year", main="DAX Index (1991-1999)", col = "blue")
abline(v=1997, col="red")

```

####c.Decompose the time series into its components (i.e., trend, seasonality, random). Keep in mind that this is a multiplicative model you want. Create a plot of all decomposed components. As before, make all lines blue and have a vertical divider at the year 1997.

```{r DAX decomposition, echo = TRUE}

# Decompose components using multiplicative methods
DAXdataComp <- decompose(DAXdata, type = c("multiplicative"))
plot(DAXdataComp, col="blue")
abline(v=1997, col="red")

```

###2. Temperature Data

####a. Using the maxtemp dataset granted by loading fpp2, there are maximum annual temperature data in Celsius. For more information, use help(maxtemp). To see what you're looking at, execute the command in 'Examples' in the help document.

```{r Read in the maxtemp dataset, run a quick plot, echo = TRUE, warning = FALSE}

# Maximum annual temperatures (degrees Celsius) for Moorabbin Airport, Melbourne. 1971-2016.
# Annual time series of class ts.
# Examples
autoplot(maxtemp)

```

####b. We are only concerned with information after 1990. Please eliminate unwanted information or subset information we care about.

```{r Clean data for post 1990 temperatures, echo = TRUE}

maxtemp_1990_2016 <- window(maxtemp,start=1990,end=2016)
plot(maxtemp, ylab="Maximum Temperature", xlab="Year",main="Maximum Temperature: Melborne, Australia", xlim=c(1990,2021))

```

####c. Utilize SES to predict the next five years of maximum temperatures in Melbourne.Plot this information, including the prior information and the forecast. Add the predicted value line across 1990-present as a separate line, preferably blue. So, to review, you should have your fit, the predicted value line overlaying it, and a forecast through 2021, all on one axis. Find the AICc of this fitted model. You will use that information later.

```{r ses fitted curve and predictions, echo = TRUE}

# SES analysis
# exponential smoothing
# simple means start with first sample, h=5 to predict out 5 years
maxtemp_ses_1990_2016_5y <- ses(maxtemp_1990_2016, alpha=0.8,beta=0.2,h=5)

# Predicted Values
plot(maxtemp_ses_1990_2016_5y, ylab="Temperature", xlab="Year", fcol="white", type="o", col="blue", main="Maximum Temperature Forecasts from Simple Exponential Smoothing")

# Fitted Values
fit_maxtemp_ses_1990_2016_5y <- fitted(maxtemp_ses_1990_2016_5y)
lines(fit_maxtemp_ses_1990_2016_5y, col="red", type="o")

# Forecasted Values
lines(maxtemp_ses_1990_2016_5y$mean, col="green", type="o")

# plot datelime from 2016
abline(v=2016, col="red")

legend("topleft", lty=1, col=c("blue", "red", "green"), c("Predicted", "Fitted", "5 Year Forecast"), pch=1)

``` 

```{r ses AICc, echo = TRUE}

# AICc of this ses model: 148.3759
maxtemp_ses_1990_2016_5y$model

```
####d.Now use a damped Holt's linear trend to also predict out five years. Make sure initial="optimal." As above, create a similar plot to 1C, but use the Holt fit instead.

```{r Holt dampened fitted curve and predictions, echo = TRUE}

# Holt analysis
# Damped linear trend 
# optimal means the initial values are optimized along with the smoothing parameters, h=5 to predict out 5 years
maxtemp_holt_1990_2016 <- holt(maxtemp_1990_2016, alpha=0.8, beta=0.2, damped=TRUE, initial='optimal', h=5)

# Predicted Values
plot(maxtemp_holt_1990_2016, ylab="Temperature", xlab="Year", fcol="white", type="o", main="Maximum Temperature Forecasts from Holt's Damped Smoothing", col="blue")

# Fitted Values
fit_holt_1990_2016 <- fitted(maxtemp_holt_1990_2016)
lines(fit_holt_1990_2016, col="red", type="o")

# The Forecasted Values
lines(maxtemp_holt_1990_2016$mean, col="green", type="o")

legend("topleft", lty=1, col=c("blue", "red", "green"), c("Predicted", "Fitted", "5 Year Forecast"), pch=1)

```
```{r Holt AICc, echo = TRUE}

# AICc of this holt model: 157.9802
maxtemp_holt_1990_2016$model

```

####e. Compare the AICc of the ses() and holt() models. Which model is better here?

#####AICc of ses model: 148.3759
#####AICc of holt model: 157.9802
#####Based on the Wikipedia entry on AICc, the smaller value AICc indicates the better fitting model.
#####Here, the ses model is the best of the two cases.

###2. The Wands Choose the Wizard

####a. Utilize the dygraphs library. Read in both Unit11TimeSeries_Ollivander and _Gregorovitch.csv as two different data frames. They do not have headers, so make sure you account for that. This is a time series of Wands sold over years.

```{r Read in csv datasets for Ollivander and Gregorovitch, echo = TRUE}

# Read in csv datasets for Ollivander and Gregorovitch
# Rename columns
Ollivander <- read.csv("./Unit11TimeSeries_Ollivander.csv", header=FALSE)
Ollivander <- dplyr::rename(Ollivander, "Year" = V1, "Wands" = V2)

Gregorovitch <- read.csv("./Unit11TimeSeries_Gregorovitch.csv", header=FALSE)
Gregorovitch <- dplyr::rename(Gregorovitch, "Year" = V1, "Wands" = V2)

```

####b. You don't have your information in the proper format! In both data sets, you'll need to first convert the date-like variable to an actual Date class.

```{r Convert "Year" to Date class variable for Ollivander and Gregorovitch, echo = TRUE}

# Convert "Year" to Date class variable for Ollivander and Gregorovitch
Ollivander$Year <- as.Date(Ollivander$Year,"%m/%d/%Y")
Gregorovitch$Year <- as.Date(Gregorovitch$Year,"%m/%d/%Y")

```

####c.Use the library xts (and the xts() function in it) to make each data frame an xts object (effectively, a time series). You'll want to order.by the Date variable.

```{r Convert Ollivander and Gregorovitch data frames into xts (time series) objects, echo = TRUE}

# Convert Ollivander and Gregorovitch data frames into xts (time series) objects 
OTS <- as.xts(Ollivander, order.by = Ollivander$Year)
OTS <- xts(Ollivander$Wands, order.by=Ollivander$Year)

GTS <- as.xts(Gregorovitch, order.by = Gregorovitch$Year)
GTS <- xts(Gregorovitch$Wands, order.by=Gregorovitch$Year)

```

####d. Bind the two xts objects together and create a dygraph from it. Utilize the help() index if you're stuck.
#####+ Give an effective title and x/y axes.
#####+ Label each Series (via dySeries) to be the appropriate wand-maker. So, one line should create a label for Ollivander and the other for Gregorovitch.
#####+ Stack this graph and modify the two lines to be different colors (and not the default ones!) Any colors are fine, but make sure they're visible and that Ollivander is a different color than Gregorovitch.
#####+ Activate a range selector and make it big enough to view.
#####+ Use dyShading to illuminate approximately when Voldemort was revived and at-large: between 1995 to 1999.
#####+ Enable Highlighting on the graph, so mousing over a line bolds it.   

```{r Bind and dygrapgh Ollivander and Gregorovitch xts objects, echo = TRUE}

# Bind both xts (time series) objects 
wandTS <-cbind(OTS, GTS)
colnames(wandTS) <- c("Ollivander", "Gregorovitch")

#dygraphs
dygraph(wandTS, main="Ollivander vs. Gregorovitch Wand Sales", ylab="Wands", xlab="Year") %>%
  dyOptions(rightGap=20) %>%
  dyLegend(width=291) %>%
  dyAxis('y', rangePad=10) %>%
  dySeries(name = "Ollivander", label = "Ollivander", color = "red") %>%
  dySeries(name = "Gregorovitch", label = "Gregorovitch", color = "darkgreen") %>%
  dyRangeSelector(height=100, fillColor = "purple") %>%
  dyShading(from = "1995-1-1", to = "1999-1-1", color = "#87f89b") %>%
  dyHighlight(highlightCircleSize = 5,
           highlightSeriesBackgroundAlpha = 0.5,
                       hideOnMouseOut = FALSE) %>%
  dyEvent("1995-1-1", "Voldemort Rises", labelLoc = "top") %>%
  dyEvent("1999-1-1", "Voldemort Defeated", labelLoc = "top")

```




