---
title: "Live Session Unit 4 Assignment"
author: "Luke Pierce"
date: "February 5, 2018"
output:
   html_document:
     keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
####GitHub Repository: https://github.com/lp5510/Homework/tree/master/LiveSession04

# Questions

### 1. FiveThirtyEight Data:

####Navigate on GitHub to https://github.com/rudeboybert/fivethirtyeight and read README.md. Seriously, it will include everything you need.

https://github.com/rudeboybert/fivethirtyeight/blob/master/README.md

####Install the fivethirtyeight package.


```{r, fivethirtyeight_Package, echo = TRUE}

#install.packages("fivethirtyeight")

```

####In the listing of Data sets in package 'fivethirtyeight,' assign the eighteenth data set to an object 'df.'

```{r, fivethirtyeight_file_download, echo = TRUE}

# Download csv file from URL

library("fivethirtyeight")
library(repmis)
library(RCurl)

URL <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/recent-grads.csv" 

download.file(URL, destfile = "./recent_grads.csv")

# Assign object to df

df <- read.csv("recent_grads.csv", header=TRUE)

```

####Use a more detailed list of the data sets to write out the URL in a comment to the related news story.

In the FiveThirtyEight article "The Economic Guide to Picking a College Major"(https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/),
Ben Casselman presents an interesting analysis of the financial value of a 4 year college degree, where the data supports the notion that the most popular degree, Psychology, pays a below median wage of $31,500, while the far more unpopular (and difficult) Petroleum Engineering major can expect initial earnings of $110,000. It would be interesting to compare other technical but non-college based career paths in such fields in the real estate, plumbing or electrical trades.

####Using R command(s), give the dimensions and column names of this data frame.

```{r, df_commands, echo = TRUE}

# Dimensions and Column Names
dim(df)
colnames(df)

```

### 2. Data Summary: Use your newly assigned data frame for Question 2. 

####Write an R command that gives you the column names of the data frame.
Right after that, write one that counts the number of columns but not rows. Hint: The number should match one of your numbers in Question 1d for dimensions. 
 
```{r, df_ncol_command, echo = TRUE}

# Column Names and number of columns
colnames(df)
ncol(df)

```
 
####Generate a count of each unique major_category in the data frame.
I recommend using libraries to help.  I have demonstrated one briefly in live-session.  To be clear, this should look like a matrix or data frame containing the major_category and the frequency it occurs in the dataset. Assign it to major_count. 

```{r, unique_data_count, echo = TRUE}

library(plyr)
major_count <- count(df, 'Major_category')
major_count

```

####Make a barplot of major_count.
To make things easier to read, put par(las=2) before your plot to make the text perpendicular to the axis.   Make sure to label the title with something informative (check the vignette if you need), label the x and y axis, and make it any color other than grey.  Assign the major_category labels to their respective bar.  Flip the barplot horizontally so that bars extend to the right, not upward.  All of these options can be done in a single pass of barplot(). Note: It's okay if it's wider than the preview pane.

```{r, Category_of_College_Majors_plot, echo = TRUE}

library(ggplot2)

ggplot(major_count, aes(x=(Major_category), y=(freq))) +
  geom_bar(stat='identity', fill = 'red') +
  coord_flip() + labs(title ="Category of Major vs. Frequency", x = "Category of Major", y = "Frequency")


```

####Write the fivethirtyeight data to a csv file.
Make sure that it does not have row labels.

```{r, write_major_count_df_to_csv, echo = TRUE}

write.csv(major_count, "major_count.csv", row.names=F)

```
